Region-Network Hierarchical Sparsity Priors for\\ High-Dimensional Inference in Brain Imaging

### 2016-02-05 paper/icml2016/example_paper.tex
ref

Brain \textit{networks} consist of a group of one or several brain regions.
The set of brain networks \(\mathcal{H}\) also forms a partition of 
\(\{1, \dots, p\}\) and in addition it is consistent with \(\mathcal G\) in
the sense that
\[\textrm{for all }g\in\mathcal G, h\in\mathcal H,\quad
\textrm{ either } g\subset h\textrm{ or }g\cap h = \emptyset.\]
This allows a clear association of each region \(g\in\mathcal G\) to a 
network \(h\in\mathcal H\) and thus establishes a tree structure (up to 
adding a root node containing all voxels).

@@

One can read off the formula that decreasing \(\alpha\) leads to less 
penalization of brain networks and thus to tendentially fully active groups
and dense brain maps. If at the same time \(\beta\) is increased to a point
of inducing group sparsity, then only the structure of brain regions encoded
by \(\mathcal G\) is present. Conversely, if \(\beta\) is decreased to 
become non-selective, and \(\alpha\) is increased to group selectivity,
then the structure imposed will come from \(\mathcal H\), leading to the
selection of brain networks rather than regions.

It is important to note that the above trade-off shows that predominance
can be either attributed to regions or networks, but that nevertheless the
penalty structure is hierarchical: If the network penalty layer sets a
network group to zero, then all the contained region groups are forced to
have activity zero. Conversely, if a brain region has non-zero coefficients,
then necessarily the network containing it must be active.
This relation is asymmetric, since when swapping the roles of \(\mathcal G\)
 and \(\mathcal H\), these statements do not hold true anymore. A brain 
region can set all its coefficients to zero without forcing the 
corresponding network to zero. A brain network can be active without its
subregions necessarily being active.
When evaluating the trade-off in \((\alpha, \beta)\), this needs to be taken
into account.

The prediction problem at hand is a multiclass classification. We choose to
attack this using one-vs-rest scheme on a binary logistic regression, whose
loss can be written as
\[\sum_{i=1}^n\log(1 + \exp(-y_i\langle x_i, w\rangle)) + \lambda\Omega(w),\]
if \(y\in\{-1, 1\}\) and with \(x_i\in\R^p\) the training sample brain images.
The data loss function is approximated using an iterative forward-backward
scheme analogous to FISTA for the lasso \cite{beck2009}.

@@

% \paragraph{Hyperparameter optimization.}
\subsection{Hyperparameter optimization}
The stratified and shuffled training data were submitted
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.

>>>>

Brain \textit{networks} consist of a group of one or several brain regions.
The set of brain networks \(\mathcal{H}\) also forms a partition of 
\(\{1, \dots, p\}\) and in addition it is consistent with \(\mathcal G\) in
the sense that
\[\textrm{for all }g\in\mathcal G, h\in\mathcal H,\quad
\textrm{ either } g\subset h\textrm{ or }g\cap h = \emptyset.\]
This allows the association of each region \(g\in\mathcal G\) to a 
network \(h\in\mathcal H\) and thus establishes a tree structure (up to 
adding a root node containing all voxels).

@@

One can read off the formula that decreasing \(\alpha\) leads to less 
penalization of brain networks and thus to tendentially fully active groups
and dense brain maps. If at the same time \(\beta\) is increased to 
induce group sparsity, then only the structure of brain regions encoded
by \(\mathcal G\) is present. Conversely, if \(\beta\) is chosen 
sufficiently small, and \(\alpha\) increased,
then the structure will come from \(\mathcal H\), leading to the
selection of brain networks rather than regions.

It is important to note that the above trade-off shows that predominance
can be either attributed to regions or networks, but that nevertheless the
penalty structure is hierarchical: If the network penalty layer sets a
network group to zero, then all the contained region groups are forced to
have activity zero. Conversely, if a brain region has non-zero coefficients,
then necessarily the network containing it must be active.
This relation is asymmetric, the roles of \(\mathcal G\) and \(\mathcal H\) 
cannot be swapped: A brain 
region can set all its coefficients to zero without forcing the 
corresponding network to zero. A brain network can be active without its
subregions being active.
When evaluating the trade-off in \((\alpha, \beta)\), this needs to be taken
into account.

The prediction problem at hand is a multiclass classification. We choose to
attack this using one-vs-rest scheme on a binary logistic regression, whose
loss can be written as
\[\sum_{i=1}^n\log(1 + \exp(-y_i\langle x_i, w\rangle)) + \lambda\Omega(w),\]
if \(y\in\{-1, 1\}\) and with \(x_i\in\R^p\) the training sample brain images.
We optimize parameters \(w\) using an iterative forward-backward
scheme analogous to FISTA for the lasso \cite{beck2009}.

@@

% \paragraph{Hyperparameter optimization.}
\subsection{Hyperparameter optimization}
The stratified and shuffled training data were submitted
to a nested cross-validation scheme
for model selection and model assessment.
In the inner CV layer, the logistic regression estimators
have been trained in a one-versus-rest design that
distinguishes each class from
the respective 17 other classes
(number of maximal iterations=$100$, tolerance=$0.001$).
{\color{red} In the outer CV layer, grid search
selected among candidates for the respective $\lambda$ parameter
by searching between $10^{-2}$ and $10^{1}$ in 9 steps on a logarithmic scale.}
Importantly, the thus selected sparse logistic regression classifier was
evaluated on an identical test set in all settings.
